### 2015-07-09 narrative/data-description.tex
short description on the auditor

%
From an application level, we will explore these techniques through our ADIOS
framework using many of our existing application partners, including
XGC1, GTC, and SPECFM3D. We will investigate: 1) what hints the users can give
the system, and 2) how the users can replace much of their
application specific code with
``system level code'' through the use of these hints.  We will design an API
that will be integrated with the storage layer to give time estimations so that
users can decide which data priorities will be written and read. The utility
value for each chunk will be used to sort the chunks into prioritized bins.
Chunks in a particular bin will be assigned an appropriate storage
tier and lifetime, which can be specified by the user through the API.
As an example, Figure~\ref{fig:ssio-bucket} illustrates a use case for the XGC1 code.
The user can pick a data-refactoring scheme (see
Sec~\ref{sec:data-refactor}) which will classify the particle data into four different
prioritization levels. The user can then pick a data utility, such as three months
for the highest priority items and 12 months for the lowest priority data.
The user will then call an API function that allows them
to specify all of the chunks they want
written (particles, two fields, and a mesh in this example) and the system
will return a time estimate for this write operation.
If this time estimate is too large, an API call can be made to 
specify the lowest priority element to be written, skipping the rest in order to save time.
The data will then be stored
across the layers of the file system where the two highest priority buckets
will be stored in the Parallel File System, but the next highest priority items
will be stored lower in the storage stack. Depending on the capabilities of the actual storage system, 
SSIO will decide if the data has to go through the parallel file system 
towards a lower stack or can be staged to that stack directly. 
% Description of the utility function research

>>>>

From an application level, we will explore these techniques through our ADIOS
framework using many of our existing application partners, including
XGC1~\cite{chang2006integrated}, GTC~\cite{klasky2003grid} , and SPECFM3D~\cite{SPECFEM3D}. We will investigate: 1) what hints the users can give
the system, and 2) how the users can replace much of their
application specific code with
``system level code'' through the use of these hints.  We will design an API
that will be integrated with the storage layer to give time estimations so that
users can decide which data priorities will be written and read. The utility
value for each chunk will be used to sort the chunks into prioritized bins.
Chunks in a particular bin will be assigned an appropriate storage
tier and lifetime, which can be specified by the user through the API.
As an example, Figure~\ref{fig:ssio-bucket} illustrates a use case for the XGC1 code.
The user can pick a data-refactoring scheme (see
Sec~\ref{sec:data-refactor}) which will classify the particle data into four different
prioritization levels. The user can then pick a data utility, such as three months
for the highest priority items and 12 months for the lowest priority data.
The user will then call an API function that allows them
to specify all of the chunks they want
written (particles, two fields, and a mesh in this example) and the system
will return a time estimate for this write operation.
If this time estimate is too large, an API call can be made to 
specify the lowest priority element to be written, skipping the rest in order to save time.
The data will then be stored
across the layers of the file system where the two highest priority buckets
will be stored in the Parallel File System, but the next highest priority items
will be stored lower in the storage stack. Depending on the capabilities of the actual storage system, 
SSIO will decide if the data has to go through the parallel file system 
towards a lower stack or can be staged to that stack directly. 
% Description of the utility function research

### 2015-07-09 narrative/data-reduction.tex
short description on the auditor

\paragraph{High Impact Challenges:} Maximal data reduction on the SSIO layer
would result from storing only the code and the input parameters, and simply
recalculating all of the results. This extreme is not feasible in practice
owing to the size of the computation involved, but sheds light on the path to
achieving a step change in data reduction. We shall investigate the possibility
of writing a minimal set of information, and placing an ``auditor'' code in a
docker container so that data can be re-generated with a {\em local}
approximation of the full simulation. This is an entirely new approach to
tackle the data overload by using preprocessors in which information is not
simply stored in the form of raw data, but rather in the form of appropriate
parameterized algorithms. A priori information on the nature of the data is
used to choose the basic algorithm while the actual values of the data are used
to tune the parameters used in the reconstruction procedures. We shall refer to
this paradigm as an Auditor.  We will examine the use of several auditors for
the simulations which we will work with. One approach is to use a fluid-code to
audit a kinetic code, which can be a good approximation over short periods of
time.

>>>>

\paragraph{High Impact Challenges:} Maximal data reduction on the SSIO layer
would result from storing only the code and the input parameters, and simply
recalculating all of the results. This extreme is not feasible in practice
owing to the size of the computation involved, but sheds light on the path to
achieving a step change in data reduction. We shall investigate the possibility
of writing a minimal set of information, and placing an ``auditor'' code in a
docker container so that data can be re-generated with a {\em local}
approximation of the full simulation. This is an entirely new approach to
tackle the data overload by using preprocessors in which information is not
simply stored in the form of raw data, but rather in the form of appropriate
parameterized algorithms. An ``auditor'' is a code which mimics the exascale simulation, 
but typically can run using less degrees of freedom. This means that the code is a good
local (in space and time) approximation of the exascale simulation and will be much less
expensive to execute (so that it can be run in situ), and follows the behavior of the simulation
for short periods of time. This allows us to ``audit'' the simulation and when the differences 
become large in a local region of space, we write out the results of the data in this local region.
The hope is that this will be done only locally and that with a good ``auditor''  this will not happen
too frequently. This then allows us to use the ``auditor'' to reproduce results during the analysis 
process using an inexpensive computation and put less pressure on the SSIO layer.
%
% 
%   A priori information on the nature of the data is
%used to choose the basic algorithm while the actual values of the data are used
%to tune the parameters used in the reconstruction procedures. We shall refer to
%this paradigm as an Auditor.  We will examine the use of several auditors for
%the simulations which we will work with. One approach is to use a fluid-code to
%audit a kinetic code, which can be a good approximation over short periods of
%time.
